### 2018-12-19 Response_to_Nature_Comms.Rmd
luke letter edits

---
title: 'Sexual selection improves population fitness: a meta-analysis'
output:
  pdf_document: default
  word_document: default
subtitle: Response to Decision Letter
bibliography: Manuscripts/Bibliography/Bibliography.bib
---

```{r, warning=FALSE, message=FALSE, results='hide', echo = FALSE}
library(pander)
library(tidyr)
library(compute.es)
library(metafor)
library(plyr)
library(dplyr)
library(lme4)
library(car)
library(forestplot)
library(ggplot2)
library(ggthemes)
library(kableExtra)
library(ggrepel)
library(reshape2)
library(RColorBrewer)
library(ggridges)
library(rstan) #Note that installation requires some effort: dependency for brms
library(brms) 
library(backports) #seems to be a dependency
library(bayesplot)
#devtools::install_github("mvuorre/brmstools")
library(brmstools)
library(metaAidR) # install.packages("metaAidR")
library(cowplot)
library(gridExtra)
library(janitor)
#setwd("~/Documents/Uni/metaanalysis/SexualSelection")
source("I2_function.R") #Adapted function for obtaining I2 with CIs
source("Vdodge_function.R") # nice function for ggplot
source("Tidy_functions_for_brms.R") #Tidy functions for making model tables for brms

load(".RData")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Dear Dr. Jones and colleagues,

Many thanks for your time, and for the very helpful comments on our manuscript. We have considerably reworked the manuscript in light of the many insightful comments and suggestions provided by the reviewers. Please see below for a detailed list of responses to each of the reviewers' points.

We also attach a revised manuscript and a revised copy of the supplementary material, as well as a copy of the revised manuscript in which the significant changes and additions have been highlighted. Please let us know if we can do anything further to assist you.

Finally, in addition to the changes suggested by reviewers, we discovered a small number of errors in our original analysis, related to how effect sizes were calculated from certain primary studies in our meta-analysis. As such, we have thoroughly rechecked our dataset and re-ran the analyses with all errors corrected. **XXXXXXXXX**

Yours Sincerely,

Justin Cally and co-authors

________________________________________________________

## Comments from Reviewer 1 (Prof. Jacek Radwan)

>*The manuscript presents results of a meta-analysis assessing the effect of sexual selection on population fitness. This topic has been increasingly studied over the past decade or so, and the time seems ripe for a synthesis. The authors have done good job searching the literature – it includes all relevant papers I could remember, and the results should be of interest to a broad range of evolutionary biologists. I cannot expertly judge on state-of-the-art methodology, but I do have several general concerns about how the analyses were performed.*

Many thanks for your time and this valuable feedback.

>*Firstly, it seems that several male traits like to be under sexual selection, but not related to population fitness were included. I don’t think this is correct given the question being asked (i.e. whether sexual selection increases population fitness). I think traits which are directly under sexual selection (eg. attractiveness, reproductive success) should not be pooled with traits which may respond to manipulation of sexual selection indirectly (eg. male development time, survival) and can affect population fitness. Distinguishing between both types of trait could actually be revealing – e.g. could expose trade-offs between sexually selected traits and fitness components unrelated to reproductive competition (see eg. Radwan et al. 2015 Evol Biol), a thus potentially explain lower effect of sexual selection on (pooled) male traits than on female traits.*

The reviewer is correct that several traits are likely under trade-offs with other traits as sexual selection not only affects mean fitness but alters life history traits (*i.e.* sexual selection may increase male body size or attractiveness, but decrease development rate). In essence this is why our meta-analysis sought to investigate the *net* effect of sexual selection on population fitness across many traits, rather than one that may be traded off in a given direction. Importantly, to account for the variable and uncertain relationships certain traits have with fitness (as raised by the reviewer). We provide a table (Table S1) on how each trait is measured and their relationship with fitness. Given that this information was in the Supplementary Material we have **revised the manuscript to clarify the distinctions more thoroughly... Probably need to check this again**.

Additionally, the reviewers comment that male attractiveness may not be related to population fitness is correct. Because male attractiveness may be positively or negatively correlated with average fitness (although in many condition-dependent traits a positive association is assumed) we have re-classified *Male Attractiveness* as ambiguously related to fitness, rather than indirect. All subsequent models, statistics and figures now take into account this re-classification. However, given the small sample size for male attractiveness effect sizes (*n* = 6), we find only marginal changes in the results and no key changes in the key findings of the study.

>*Secondly, looking at Figure 1, one notices apparent anomalies, for example significantly positive slope for male attractiveness, based on studies with average effect size close to zero. I think this (and few others) surprising estimates may result from fitting random terms across all trait types (fitted as a fixed factor, second model); I guess fitting interaction (random slopes) would not be feasible for some categories including few data points, but some of them could easily be pooled in wider categories.*

The reviewer raises questions of the analysis that produce seemingly suprising results. We agree that some of *significant* predictions in Figure 1 (whereby 95 % CIs do not overlap zero) appear surprising based on the distribution on effect sizes. In regards to the trait "male attractiveness", the 95 % CIs do not overlap zero. This is the case even though the higher weighted studies are very close to zero. However, this seeming anomoly can be explained by various reasons.

Firstly, although the effect size prediction for the trait has 95 % CIs not overlapping zero, the overall magnitude of the effect size is small (see Table exerpt), and the Bayes Factor (an estimate of the likelihood that the effect size is actually positive) is also small. The full table for all predicted effect sizes used in Figure S1 can be seen in **S8**. Below we provide the exerpt of this table and forest plot for "Male attractiveness".

Secondly, the effect size predictions were made using a single model that incorperated all 459 effect sizes, thus model predictions for each trait have narrower confidence intervals as the fitted values do not incorperate measurement error. In regards to our R code and analysis, this means that the predictions that appeared in Figure 1 are sourced from using ``fitted.brmsfit`` as opposed to ``predict.brmsfit``. In order to clarify this seeming anomaly in model predictions, some meta-analyses investigating the effect of a variable on more than one outcome model each outcome (in this case the *fitness component*) separately. Arguably, this method provides more variable model estimates as models will always have lower power due to reduced sample sizes and the incorporation of measurement error in the predictions. We have added a table to the Supplementary Material where models are independently conducted on all fitness components where n > 3 (**Table S9**). This may be of particular value for researchers using our meta-analysis as a research tool, whereby they are interested in the effect sizes from only a single trait (as opposed to our focus on population fitness as a whole).

Given, the confusion with regards to Figure 1 as it originally appeared within the manuscript we have altered the figure to a generalized forest plot by grouping the effect sizes from each of the fitness classes (ambiguous, indirect and direct). The previous Figure 1 has now been moved to the Supplementary Material with corresponding tables.

```{r}
left_join(fitness.component.predictions, predictions.outcomes, by = "Fitness Component")  %>% filter(`Fitness Component` == "Male Attractiveness") %>% pander(split.table = Inf, digits = 2)
```

```{r, fig.height=2.5, fig.width=6}
pd <- position_dodgev(0.6)

p.meta.male.attract <- full_dataset %>% 
  mutate(Sex = replace(as.character(Sex), Sex == "B", "Both"),
         Sex = replace(Sex, Sex == "M", "Male"),
         Sex = replace(Sex, Sex == "F", "Female"),
         Outcome.Class = factor(Outcome.Class, levels = c("Ambiguous", "Indirect", "Direct"))) %>%
  filter(Outcome == "Male Attractiveness") %>%

  ggplot(aes(x=reorder(AuthorYear, -g), y = g)) +
  scale_color_manual(values = c("Ambiguous" = "#a50f15", "Indirect" = "#fe9929", "Direct" = "#4daf4a"), 
                     name = "Relationship\nto fitness")+
  scale_shape_manual(values=c(21,22,24))+
  scale_fill_manual(values = c("Ambiguous" = "#a50f15", "Indirect" = "#fe9929", "Direct" = "#4daf4a"), 
                    name = "Relationship\nto fitness")+
  geom_linerange(aes(ymin = lowerci, 
                   ymax = upperci,
                   x = reorder(AuthorYear, -g),
                     color = Outcome.Class,
                     group = "identity"),
               show.legend = FALSE,
               position = position_dodge2(width = 0.5, preserve = "single")) +

  geom_point(aes(shape = Sex,
                 fill = Outcome.Class),
             size = 1.75, 
             color = "grey20",
             position = position_dodge2(width = 0.5),
             show.legend = F) +

  coord_flip()+

  scale_y_continuous(limits=c(-3.1, 3.1), 
                     breaks = c(-3, -2, -1, 0, 1, 2, 3), 
                     name='Standardized Mean Difference (g) \n[positive values indicate sexual selection improves fitness components]') +

  xlab('Reference') +

  geom_hline(yintercept=0, 
             color='black', 
             linetype='dashed')+

    geom_hline(yintercept=0.314, 
             color='#4daf4a', 
             linetype=1,
             size = 1)+

  geom_hline(yintercept=0.0302, 
             color='#4daf4a', 
             linetype=3)+

  geom_hline(yintercept=0.61, 
             color='#4daf4a', 
             linetype=3)+

  facet_grid(Outcome_f~.,
             labeller = label_wrap_gen(width=10),
             scales= 'free', 
             space='free')+

  guides(fill = guide_legend(override.aes = list(shape = 21, colour = "grey20", size = 6)),
         shape = F)+

  #Add theme specifying text size, margins, lines etc.
  theme_bw()+

  theme(strip.text.y = element_text(angle = 0, size = 12, margin = margin(t=15, r=15, b=15, l=15)), 
        strip.background = element_rect(colour = NULL,
                                        linetype = "blank",
                                        fill = "gray90"),
        text = element_text(size=13),
        panel.spacing = unit(0.5, "lines"),
        panel.border= element_blank(),
        axis.line=element_line(), 
        panel.grid.major.x = element_line(linetype = "solid", colour = "gray95"),
        panel.grid.major.y = element_line(linetype = "solid", color = "gray95"),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(), 
        legend.text = element_text(size=12), 
        legend.title=element_text(size=12, 
                                  face = "bold"),
        axis.title.x = element_text(hjust = 0.5, size = 12))

p.meta.male.attract
```

When the effect sizes (n = 6) for male attractiveness are viewed isolated from others, one can see that the large effect size (Nelson 2013) shifts the overall predicted effect size to be marginally positive.

>*Thirdly, type of trait measured explained 35% variance, but the authors do not explore this any further. However, examination of Fig. 1 suggest that some indirect/ambiguous fitness measures account for much of this heterogeneity and they generally have higher average effect sizes that direct measures (except for immunity). I’d like to see if the authors recover their main result if they only direct measures.*

**Add more in discussion about the variability amongst traits and why some might be subject to trade offs (e.g. immunity): We do kind of talk about this and an earlier draft had some of Jaceks work with regards to sexual conflict maintaing genetic variance through trade-offs. Should we add this back in?**

The reviewer provides an accurate shortcoming of discussion on the observed heterogeneity between fitness components. The sixth paragraph of our discussion is raises some potential reasons why heterogeneity may be high within our meta-analysis. Additionally, it may please the reviewer to find an addition to our Supplementary Material (Table S9), now conducts independent meta-analyses and obtains $I^2$ for each of these traits. Even when metanalyses are conducted on a single trait, heterogeneity ($I^2$) is still high > 90 %.

In regards to conducting a meta-analysis on just direct fitness components, the reviewer provides a valid question of our use of direct **and** indirect fitness components for the meta-analysis investigating the effects of sex and stress. Based on the question we asked we were interested in the effect of sexual selection on many components of fitness.

Nevertheless, we conducted our analysis using just direct fitness measures. We find the same key associations; that sexual selection generally elevates fitness, especially for females, with the effect magnified in stressful environments. Below we present the findings. 
**should we put this in the supp?**

Additionally, in order to further distinguish the effects of sexual selection on ambiguous, indirect and direct fitness components, we have structured the second paragraph of the results to more consistently reflect our classification system.

```{r, warning=FALSE, message=FALSE, error = FALSE, fig.align='center', fig.width = 10, fig.height=6}
model.direct.only <- rma.mv(g, V = var.g, 
                          mods = ~ 1 + Sex * Environment, 
                          random = list(~ 1 | Study.ID, 
                                       ~ 1 | Outcome,
                                       ~ 1 | Taxon), 
                          method = "REML", 
                          data = strict_dataset %>% filter(Outcome.Class == "Direct"))

#Generate predictions without taxon utilising the previously described function
get.predictions.direct <- function(newdata){
  B<-0; F<-0; Stressed<-0; interaction1<-0; interaction2<-0; interaction3<-0
  if(newdata[1] == "B") B<-1 
  if(newdata[1] == "F") F<-1 
  if(newdata[2] == "Stressed") Stressed<-1
  if(newdata[1] == "B" & newdata[2] == "Stressed") interaction1<-1
  if(newdata[1] == "F" & newdata[2] == "Stressed") interaction2<-1

  predict(model.direct.only, newmods=c(B, F, Stressed, interaction1=interaction1, interaction2=interaction2))
}

# Get the predictions for each combination of moderators
predictions.direct <- as.data.frame(expand.grid(Sex = c("M", "B", "F"),
                           Environment = c("Unstressed", "Stressed")))
predictions.direct <- cbind(predictions.direct, do.call("rbind", apply(predictions.direct, 1, get.predictions.direct))) %>%
  select(Sex, Environment, pred, se, ci.lb, ci.ub) 
for(i in 3:6) predictions.direct[,i] <- unlist(predictions.direct[,i])

countpred <- count_(strict_dataset %>% filter(Outcome.Class == "Direct"), c("Sex", "Environment"))

predictions.direct <- left_join(predictions.direct, countpred, by = c("Sex", "Environment"))
colnames(predictions.direct) <- c("Sex", "Environment", "Prediction", "SE", "CI.lb", "CI.ub", "n")
predictions.direct <- predictions.direct %>%
      mutate(Sex = replace(as.character(Sex), Sex == "B", "Both"),
         Sex = replace(Sex, Sex == "M", "Male"),
         Sex = replace(Sex, Sex == "F", "Female"),
         Environment = replace(as.character(Environment), Environment == "Stressed", "Stressful"),
         Environment = replace(Environment, Environment == "Unstressed", "Benign"),
         Sex = factor(Sex, levels = c("Male", "Both", "Female")))

predictions.direct <- format(predictions.direct, digits = 2)
predictions.direct$Prediction = as.numeric(predictions.direct$Prediction)
predictions.direct$CI.lb = as.numeric(predictions.direct$CI.lb)
predictions.direct$CI.ub = as.numeric(predictions.direct$CI.ub)
predictions.direct$n = as.numeric(predictions.direct$n)

if(!file.exists("data/brms.direct.rds")){
  brms.direct <- brm(g | se(SE)  ~ 1 + Sex * Environment 
                        + (1|Taxon) 
                        + (1|Study.ID) 
                        + (1|Outcome), 
                        family = "gaussian", 
                        seed = 1,
                        cores = 4, chains = 4, iter = 4000, 
                        control = list(adapt_delta = 0.9999, max_treedepth = 15),
                        data = strict_dataset %>% 
                          filter(Outcome.Class == "Direct") %>% 
                          mutate(SE = sqrt(var.g)))
  saveRDS(brms.direct, file = "data/brms.direct.rds")
}
brms.direct <- readRDS(file = "data/brms.direct.rds")

# lternatively you can obtain posterior samples manually.
post.direct <- (posterior_samples(brms.direct, 
                           pars = c("b_Intercept", "b_SexB", "b_SexF", 
                                    "b_EnvironmentStressed", "b_SexB:EnvironmentStressed", 
                                    "b_SexF:EnvironmentStressed")) %>%
           mutate(both_benign = b_Intercept + b_SexB,
                  both_stressful = b_Intercept + b_SexB + b_EnvironmentStressed + `b_SexB:EnvironmentStressed`,
                  male_benign = b_Intercept,
                  male_stressful = b_Intercept + b_EnvironmentStressed,
                  female_benign = b_Intercept + b_SexF,
                  female_stressful = b_Intercept + b_SexF + b_EnvironmentStressed + `b_SexF:EnvironmentStressed`))[,-(1:6)]

# Add columns for Environment and Sex
post.direct <- as.data.frame(t(post.direct))
post.direct$Sex <- c("Both", "Both", "Male", "Male", "Female", "Female")
post.direct$Environment <- c("Benign", "Stressful", "Benign", "Stressful", "Benign", "Stressful")

#Clean up dataframe
post.direct <- melt(post.direct, id = c("Sex", "Environment"))
post.direct$variable <- NULL

#Plot the posterior values from the Bayesian model as density ridges
pd <- position_dodgev(height = 0.3)
posterior.direct.plot <- post.direct %>% mutate(Sex = factor(Sex, levels = c("Male", "Both", "Female"))) %>% ggplot()+
  stat_density_ridges(aes(x=value, y = Environment, fill = Sex), alpha = 0.65, scale = 0.6, position = position_nudge(y = 0.15), height = 10, show.legend = F, quantile_lines = T, quantiles = 2)+
  geom_vline(xintercept = 0, linetype = 2, colour = "black") + 
  ylab("Environment\n")+
  xlab("\nEffect Size (Hedges' g)")+
  # scale_fill_manual(values = c("Male" = "#e41a1c", "Female" = "#377eb8", "Both" = "#4daf4a"))+
  # scale_color_manual(values = c("Male" = "#e41a1c", "Female" = "#377eb8", "Both" = "#4daf4a"))+
  scale_fill_manual(values = c("Male" = "#ff7f00", "Female" = "#984ea3", "Both" = "#4daf4a"))+
  scale_color_manual(values = c("Male" = "#ff7f00", "Female" = "#984ea3", "Both" = "#4daf4a"))+
  scale_x_continuous(limits = c(-0.75, 1.5), breaks = c(-1, -.5, 0, 0.5, 1, 1.5))+

  theme_bw()+

  theme(panel.spacing = unit(0.1, "lines"),
        text = element_text(size=16),
        panel.border= element_blank(),
        axis.line=element_line(), 
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(), 
        legend.text = element_text(size=16), 
        legend.title=element_text(size=16, face = "bold"),
        axis.title.x = element_text(hjust = 0.5, size = 14),
        axis.title.y = element_text(size = 16, hjust = 0.35, margin = margin(r=-10)),
        axis.text.y = element_text(angle = 0),
        plot.title = element_text(size = 16))

#Add the REML predictions as circles with error bars
both.direct.plots <- posterior.direct.plot + 
  geom_errorbarh(data = predictions.direct %>% mutate(Sex = factor(Sex, levels = c("Male", "Both", "Female"))), 
                 aes(xmin = predictions.direct$CI.lb,
                     xmax = predictions.direct$CI.ub, y = Environment,
                     color = Sex), 
                 height = 0, show.legend = F, position = pd)+

  geom_point(data = predictions.direct %>% mutate(Sex = factor(Sex, levels = c("Male", "Both", "Female"))), 
             aes(x = Prediction, y = Environment, size=n, fill = Sex), 
             shape=21, color = "grey20", position = pd) +

    guides(fill = guide_legend(reverse=T, override.aes = list(size = 7.5)))+
    scale_size(guide = 'none')+

  scale_y_discrete(expand=c(0.075, 0))

grid.arrange(both.plots + 
               guides(fill = FALSE)+
               ggtitle("Indirect And Direct\n")+ theme(plot.title = element_text(size = 18, face = "bold", colour = "Red")),
             both.direct.plots + ggtitle("Direct Only\n")+
               theme(axis.title.y = element_blank(), 
                     plot.title = element_text(size = 18, face = "bold", colour = "Red")), 
             nrow = 1, 
             widths = c(3,4))
```

```{r, warning=FALSE}
predictions.direct %>% pander(caption = "Model using only direct measures of fitness")
```

>*Another major problem I have with the manuscript concerns interpretation of the very intriguing finding that the response to manipulation of sexual selection was stronger for female traits compared to male traits. I’m confused by the authors’ explanation: do they assume sexual selection acted directly on females, and not only indirectly, via males? Only then things like mother to daughter heritability, or hard selection on females, should matter.*

Leave this for Luke

Good suggestion!

>*Perhaps the effect on females is indeed direct, and results from stress imposed by polygamous treatment, which magnifies direct, hard selection on females? This would be an important finding, and perhaps the authors could test it with their dataset by contrasting middle-class-neighborhood-like studies from those which allowed for female evolution. But if correct, this explanation is not exactly the effect of sexual selection, but rather enhanced selection of females due to enhanced (male induced) stress, so the interpretation of results should change.*

Not too sure what he means.

Answer

**Other comments:**

>*l. 30 – reviews on sexual conflict are OK to cite here, but there are empirical papers actually demonstrating correlation between male sexual selected traits (Harano et al. 2011; Plesnar et al. 2014) which should also be cited.*

Thank you, we have added citations of those papers as suggested.

>*l. 151 – the authors discuss beneficial effects of sexual selection on direct fitness measures such as reproductive success or offspring viability, but estimates for both of these measures actually overlapped zero! Perhaps joint analysis of direct fitness measures, as I suggested above, could support this conclusion, but currently this is an overstatement.*

Thank you for catching this -- we have tempered our conclusions in the relevant section of the Discussion, since the results provide only moderate evidence for a positive effect of sexual selection on these traits.

>*In the discussion the authors say they included the number of experimental evolution generations, but I could not find this information in methods.*

This information could be found in the supplementary material in the section *Meta-analyses including many moderator variables* as well as in Figure S12 and S13. Generations (specifically, log-Generations) had minimal effect on the effect sizes and so was removed for the "minimum adequate models" that we presented in the main text. **We now clarify this in the main text JUSTIN, MAYBE REFERENCE THOSE FIGURES WHEN SAYING IT'S NON-SIG OR SOMETHING LIKE THAT**.

>*Fig. S1 is not referred to in the main texts, is it different from Fig. 2, except that the latter contains predicted average values for fitness components?*

Our supplementary material does indeed contain all of our main figures, in order to show the data and code used to generate them. We have now made sure that e.g. Figure 2 in the HTML supplement is correctly labelled as Figure 2, to make it clear that it is not a different figure. **A JOB FOR JUSTIN HERE**

________________________________________________________

## Comments from Reviewer 2

>*This meta-analysis investigated the consequences of sexual selection experiments on trait mean and variance (comparing sexually selected groups vs control groups). Overall, the authors observe sexual selection usually increase mean and reduce variance (especially in females). The authors conclude sexual selection's benefits outweigh its detrimental effects. This meta-analysis is extremely well conducted (the use of both likelihood-based and Bayesian models for robustness), and although I am not an expert on this topic, I really enjoyed reading it, and it was very clear. Especially, I am impressed with the detailed supplement which showed the code and analysis. However, I have several comments regarding their analysis, which will increase the robustness of results and thus conclusions.*

Many thanks for your time and attention, and for the valuable feedback.

>*1 --- The use of Hedges' $g$. I understand that Hedges' $g$ was probably used because it can take interval measurements as well as ratio measurements (lnCVR can only take ratio measurements). However, $g$ cannot really deal well with heterogeneity between two groups (i.e. experimental and control groups having different variances). This is why Hedges' $g$ (Cohen's $d$) was criticized earlier.*

>Osenberg, C. W., O. Sarnelle, and S. D. Cooper. 1997. Effect size in ecological experiments: the application of biological models in meta-analysis. \textit{American Naturalist} 150:798-812.

>*As a response, they come up with log response ratio (lnRR) - see*

>Hedges, L. V., J. Gurevitch, and P. S. Curtis. 1999. The meta-analysis of response ratios in experimental ecology. \textit{Ecology} 80:1150-1156.

>*I recommend that the authors use lnRR for their effect size for mean comparison as well as Hedge's $g$ to see the robustness of their conclusions for the mean.*

Reviewer 2 is correct that Hedges' *g* is imperfect, and so we extended our meta-analysis to include two alternative measures of effect size. Firstly, we present the log response ratio (*lnRR*) as suggested by this reviewer. Secondly, we present a modified version of Hedges' *g* that attempts to account for this heterogeneity issue [@bonett2009meta], named SMDH (standardized mean difference with heteroscedastic population variances in the two groups). There are several key findings from our new analyses using lnRR and SMDH:

*Firstly*, Hedges' *g* and SMDH are highly correlated (in means and variance). This suggests that unequal variance between the treatment and control groups (heteroscedasticity) in the primary studies we examined is probably too small to greatly influence Hedges' *g*, which is interesting for the interpretation of earlier meta-analyses using Hedges' *g*.

```{r, warning=FALSE, message = FALSE, error=FALSE}
#This function writes in the correlation value
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(-2.5, 2.5, -2.5, 2.5))
  r <- cor(x, y, use="pairwise.complete.obs")
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  txt <- paste0(prefix, txt)
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0, 0, txt, cex = cex.cor * 2*r)
}

es.correlation.df <- lnRR.data %>% left_join(SMD.data %>% rename(`Hedge's g` = SMD)) %>% left_join(SMDH.data)

pairs(es.correlation.df %>% select(`Hedge's g`, lnRR, SMDH), lower.panel=panel.smooth, upper.panel=panel.cor)
# pairs(es.correlation.df %>% select(SMD_var, SMDH_var), lower.panel=panel.smooth, upper.panel=panel.cor)
```

*Secondly*, we note that lnRR and SMDH can only be calculated in cases where the primary study reported means, standard deviations and sample sizes (352 out of 459). This is a limitation, as we must discard roughly 25% of the data in order to use lnRR and SMDH, and so we think it is worthwhile to focus primarily on Hedges' *g* despite its minor limitations in the main manuscript, and use lnRR and SMDH for sensitivity analysis (to verify that the results are qualitatively the same as those produced by Hedges' *g*).

```{r, warning=FALSE, message = FALSE, error=FALSE}
predictions.complete2 %>% select(Sex, Environment, `Hedge's g sample size` = n) %>% 
  left_join(predictions.lnRR %>% select(Sex, Environment, `lnRR sample size` = n), by = c("Sex", "Environment")) %>%
  adorn_totals("row") %>% pander()
```

*Thirdly*, if we conduct a meta-analysis using lnRR as the effect size we obtain very similar results with all our main conclusions unchanged, keeping in mind that lnRR and Hedge's are on different scales (see the revised Supplementary Material). Specifically,
+ The effect of sexual selection on population fitness is generally positive.
+ This effect is magnified for females in stressful environments with a significantly positive interaction term (stress*females).

```{r, warning=FALSE, message=FALSE, error = FALSE, fig.align='center', fig.width = 10, fig.height=6}
grid.arrange(both.plots + 
               guides(fill = FALSE) + 
               ggtitle("Comparison between Hedge's g \nand lnRR effect sizes \n"), 
             lnRR.plots + ggtitle("\n\n") + xlim(-0.5,0.75), nrow = 1, widths = c(3,4))
```

*Fourthly*, the distribution of the lnRR effect sizes is non-normal (notably moreso than the distribution of Hedges' *g*), perhaps because one can get extreme values when taking a ratio (e.g. when the denominator is close to zero). Accordingly, @Lajeunesse_2015 showed that *lnRR* is problematic when quantifying the outcome of primary studies with small sample sizes, and can yield unsuitable variance estimates when the parameter scale of the study are close to zero. So, Hedges' *g* has the possible practical advantage of increasing model fit in our meta-analysis.

```{r}
grid.arrange(ggplot(data = lnRR.data, aes(x = lnRR))+
  geom_density(),

ggplot(lnRR.data, aes(sample = lnRR)) + 
  stat_qq() + 
  stat_qq_line() +
  labs(x = 'theoretical quantiles', y = "lnRR sample"),

ggplot(data = full_dataset, aes(x = g))+
  geom_density()+
  xlab("Hedges' g"),

ggplot(full_dataset, aes(sample = g)) + 
  stat_qq() + 
  stat_qq_line() +
  labs(x = 'theoretical quantiles', y = "Hedges' g sample"))
```

>*2 --- The authors may consider also doing another set of meta-analyses using lnVR (log variability ratio - proposed in Nakagawa et al. 2015).*

>Nakagawa, S., R. Poulin, K. Mengersen, K. Reinhold, L. Engqvist, M. Lagisz, and A. M. Senior. 2015. Meta-analysis of variation: ecological and evolutionary applications and beyond. \textit{Methods in Ecology and Evolution} 6:143-152.

>*As one can see in Figure 2, the mean results (Hedges' $g$) are a mirror image of the variance results (this makes sense CV controls for means). I would like to see what the absolute change in variances. Probably the authors can put the analysis using lnVR in the supplement. The results of this analysis can be discussed. Also, the mean-variance relationship between mean and variance (sd) should be verified (e.g. plot log mean and log sd or log variance).*

Many thanks -- we have now also present a meta-analysis of variation using lnVR in addition to one using lnCVR. We have the following comments on this new analysis:

*Firstly*, as the reviewer suggested, we now justify our use of lnCVR as opposed to lnVR, by assessing the mean-variance relationship in our data. The major limitation of lnVR is that it does not account for the mean-variance relationship; indeed this is why lnCVR was suggested as the favourable correction [@RN177]. We indeed found a strong positive correlation between mean and variance, suggesting our original use of lnCVR was justified.

```{r, warning=FALSE, message=FALSE, error = FALSE, fig.align='center', fig.width = 8, fig.height=3.75}
grid.arrange(full_dataset %>% ggplot(aes(x = mean.low, y = sd.low))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  xlab("Mean of Control Group")+
  ylab("Standard Deviation of Control Group") +
  geom_smooth(method = "lm", colour = "red"),

  full_dataset %>% ggplot(aes(x = mean.high, y = sd.high))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  xlab("Mean of Treatment Group")+
  ylab("Standard Deviation of Treatment Group") +
  geom_smooth(method = "lm", colour = "red"), nrow = 1)
```

*Secondly*, We conducted new Bayesian and REML meta-analyses for *lnVR*, which we present in the supplementary material and briefly discuss in the manuscript. We found a similar trend for the meta-analysis of variance using *lnCVR* and the meta-analysis of variance using *lnVR*, These are:

+  In stressful environments sexual selection leads to a decrease in the variance for females, relative to benign environments.  
+  Sexual selection reduces variance in females compared to males.

However, when using *lnVR*, the variability of fitness traits for males under sexual selection increased, *lnVR* for females in stressful environments was not statistically different from zero, and sexual selection in benign environments increased phenotypic variance for females. Model results and predictions (from the REML model) can be cound within the Supplementary Material.

```{r, warning=FALSE, message=FALSE, error = FALSE, fig.align='center', fig.width = 10, fig.height=6}
grid.arrange(both.var.plots + 
               guides(fill = FALSE) + 
               ggtitle("Comparison between lnCVR \nand lnVR effect sizes \n"), 
             both.lnVR.plots + ggtitle("\n\n"), nrow = 1, widths = c(3,4))
```

>*3 --- $I^2$ needs to be explained. $I^2$ is proposed originally here:*

>Higgins, J., and S. Thompson. 2002. Quantifying heterogeneity in a meta-analysis. \textit{Statistics in Medicine} 21:1539 - 1558.

>*But later expanded in here for mixed models (hierarchical models)*

>Nakagawa, S., and E. S. A. Santos. 2012. Methodological issues and advances in biological meta-analysis. \textit{Evolutionary Ecology} 26:1253-1274.

>*This is what the authors use. Also, it will be good to put the degree in I^2 in the context. To do this see this paper:*

>Senior, A. M., C. E. Grueber, T. Kamiya, M. Lagisz, K. O'Dwyer, E. S. A. Santos, and S. Nakagawa. 2016. Heterogeneity in ecological and evolutionary meta-analyses: its magnitude and implications. \textit{Ecology} 97:3293-3299.

Many thanks for this feedback -- we have incorporated all the suggestions within a paragraph in the Methods section.

>*4 --- Publication tests have been conducted on, I think, "meta-analytic" residuals results as suggested in Nakagawa and Santos 2012. One could use such residuals to conduct the trim and fill method and see how much the mean could move (see Nakagawa and Santos 2012). One should remember the funnel asymmetry could be caused by the presece of heterogeniety. See:*

>Egger, M., G. Smith, M. Schneider, and C. Minder. 1997. Bias in metaanalysis detected by a simple, graphical test. \textit{Br Med J} 315:629 - 634.

Thanks -- we now remind the reader that funnel plot asymmetry is not decisive evidence for publicaiton bias, because it can also result from unexplained heterogeneity in the original effect sizes. We have elected not to implement trim-and-fill since the method seems to have fallen into disfavour among many readers, and because we are unsure how to correctly implement it using the REML- and Bayesian methods we used. NB that there is no method for trim-and-fill in mixed meta-analysis in the popular R package `metafor` (which we used), suggesting the method is not simple to implement.

>*5 --- Figure 3 - are these grey envelopes 95% CI?*

Yes, and we have now included this information in the figure caption. Thank you for highlighting this.

>*6 --- the title - do the authors include "a systematic review" - so "a systematic reveiw and meta-analysis" - these two things are diferent - see:*

>Nakagawa, S., and R. Poulin. 2012. Meta-analytic insights into evolutionary ecology: an introduction and synthesis. \textit{Evolutionary Ecology} 26:1085-1099.

We think that the scope of a true systematic review (in the sense used by e.g. Cochrane reviews) is larger than our study entails. However, we did use a systematic literature search strategy and PRISMA framework of inclusion/exclusion. With this in mind, we have left our paper's title as "meta-analysis" rather than calling it a systematic review.

>*Hope my comments are useful.*

They were extremely useful, many thanks!

________________________________________________________

## Comments from Reviewer 3

**This manuscript addresses the question of whether, on average, sexual selection has a net beneficial or detrimental effect on population fitness, using a meta-analysis of experimental evolution studies comparing the fitness of populations under different intensities of sexual selection. The analyses consider the effects of sexual selection on both the mean and variance of population fitness, and test whether these effects differ for stressful vs. benign environments, or depend on the measure of population fitness.**

**These are issues of longstanding interest, and as there is now quite a wealth of experimental evolution studies addressing these questions a meta-analysis synthesising their findings is timely.**

Many thanks for your time and attention, and for the valuable feedback.

**The results of this meta-analysis broadly concur with current theoretical predictions, and are likely to interest a wide readership. Across all studies, environment types and fitness measures there was a small positive effect of sexual selection on mean population fitness, although effect sizes varied with the fitness measure used. The positive effect of sexual selection on fitness was strongest for females in stressful environments, who also showed reduced variance in fitness under sexual selection.**

**The authors discuss these findings thoughtfully, although they focus more on the (weaker) effects found for ‘direct’ measures of fitness while somewhat neglecting the effects seen for ‘indirect’ fitness measures (e.g. of attractiveness, mating latency, lifespan, ejaculate traits). These potentially deserve more consideration, especially as many appear to be male-limited traits or measured mainly in males, that would appear to have a direct bearing on male mating success (if an indirect link to overall fitness), yet despite this the overall effect of sexual selection on male fitness is not significant.**

Need to answer this in a similar line to Jaceks comments.

**I have only a few more specific comments:**

**Does your approach consider variation in the intensity of sexual selection exclusively on males? This is worth noting (e.g. statements such as that on line 98-99 might be amended to “Sexual selection on males significantly improved female fitness.”)**

Answer

**Experiments manipulating the sex ratio to alter sexual selection might simultaneously decrease sexual selection on males while increasing sexual selection on females. Would this be classified as reduced sexual selection in your analyses? I don’t think you included this aspect of study design as a moderator variable in any of your analyses – given that close to half of the effects you include come from “alternative manipulations” (line 70) perhaps you could test whether this affects effect size?**

Answer

**It seems possible that sexual selection on females, and not only on males, could affect population fitness – potentially even more directly than sexual selection on males. And might differing extents of sexual selection in each sex interact with the differences you saw in fitness measured in males vs. females?**

Answer

**L111-118 It seems a little odd that there is no residual heterogeneity in your $I^2$ estimates. From the supplemental information it is not entirely clear to me how you have adapted the function that is under development in “metaAidR”, but can you double check that you have appropriately incorporated the residual variance into these estimates?**

Answer

**L307-311 Please clarify here that lnCVR was calculated as the ratio $ln(CVfitness_{SS}/CVfitness_{no SS})$**

Answer

**Fig. 3 Please explain the dashed red and black lines in the figure caption.**

Answer

**Tables 1, 2 Please check – the parameter estimates for ‘Female sex’ and test for ‘Female > Male’ are bolded inconsistently between the REML and Bayesian models where these estimates are identical.**

Answer

# References
<div id="refs"></div>